Generalizing the Model View Controller Paradigm

### 2008-02-27 inst/paper/paper2.tex
Replaced reference to soon to be defunct class exprSet with ExpressionSet.
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MVCClass@30240 bc3139a8-67e5-0310-9ffc-ced21a209358

We encapsulate the particular data model in the class definition, which allows
for efficient dispatch, and accessor functions that are appropriate for the
underlying implementation.  The specific model classes used in the
\Rpackage{iSNetwork} package to represent data are \Rclass{dfModel},
\Rclass{graphModel}, and \Rclass{exprModel}.  The \Rclass{dfModel} class
represents a model where the \Rslot{modelData} slot has class data frame (or
matrix), the \Rclass{graphModel} class represents a model where the
\Rslot{modelData} slot has class graph, and the \Rclass{exprModel} class
represents a model where the \Rslot{modelData} slot has class exprSet, a class
used to represent gene expression data \cite{BioC}.

@@

The first message class that is discussed is
\Rclass{gAddChildMessage}.  A \Robject{gAddChildMessage} object is very
similar to a \Robject{gAddDataMessage} object because they are both adding a
new model and a new \Robject{MVC} object.  The difference with a
\Robject{gAddChildMessage} object is that it must fill in some extra
information to tie this new \Robject{MVC} object to its parent \Robject{MVC}
object.  The \Rclass{gAddChildMessage} class inherits from the
\Rclass{gAddMessage} class, with the same slots of \Rslot{dataName},
\Rslot{mData}, and \Rslot{type}.  The \Rslot{dataName} slot is the name of the
new model (and new \Robject{MVC}), the \Rslot{mData} slot is a list that
contains the
model data and virtual data to fill the \Rslot{modelData} and
\Rslot{virtualData} slots, respectively, of the new model, and the
\Rslot{type} slot gives the type of the model (currently, the options are
``exprSet'', ``graph'', or ``data.frame'').  As with the previous
message classes, the \Rclass{gAddChildMessage} class has two methods:
\Rfunction{initialize} and \Rfunction{handleMessage}.  The
\Rfunction{initialize} method properly fills the slots of the object when it
is created and the \Rfunction{handleMessage} method ensures that the new
model and \Robject{MVC} objects are created (which is what the
\Rfunction{handleMessage} method for the \Rclass{gAddDataMessage} class does)
and then it fills the \Rslot{linkData} slot of the new child model
so that the \Rfunction{toParent} and \Rfunction{fromParent} functions are
available for messaging between \Robject{MVC} objects and finally it stores
information in the child \Robject{MVC} object's controller that determines if
the child model was created from a subset of the parent
model.  Thus, the \Rfunction{handleMessage} method for the
\Rclass{gAddChildMessage} class both creates the child \Robject{MVC} object
and relates it to the parent \Robject{MVC} object.

>>>>

We encapsulate the particular data model in the class definition, which allows
for efficient dispatch, and accessor functions that are appropriate for the
underlying implementation.  The specific model classes used in the
\Rpackage{iSNetwork} package to represent data are \Rclass{dfModel},
\Rclass{graphModel}, and \Rclass{exprModel}.  The \Rclass{dfModel} class
represents a model where the \Rslot{modelData} slot has class data frame (or
matrix), the \Rclass{graphModel} class represents a model where the
\Rslot{modelData} slot has class graph, and the \Rclass{exprModel} class
represents a model where the \Rslot{modelData} slot has class ExpressionSet,
a class used to represent gene expression data \cite{BioC}.

@@

The first message class that is discussed is
\Rclass{gAddChildMessage}.  A \Robject{gAddChildMessage} object is very
similar to a \Robject{gAddDataMessage} object because they are both adding a
new model and a new \Robject{MVC} object.  The difference with a
\Robject{gAddChildMessage} object is that it must fill in some extra
information to tie this new \Robject{MVC} object to its parent \Robject{MVC}
object.  The \Rclass{gAddChildMessage} class inherits from the
\Rclass{gAddMessage} class, with the same slots of \Rslot{dataName},
\Rslot{mData}, and \Rslot{type}.  The \Rslot{dataName} slot is the name of the
new model (and new \Robject{MVC}), the \Rslot{mData} slot is a list that
contains the
model data and virtual data to fill the \Rslot{modelData} and
\Rslot{virtualData} slots, respectively, of the new model, and the
\Rslot{type} slot gives the type of the model (currently, the options are
``ExpressionSet'', ``graph'', or ``data.frame'').  As with the previous
message classes, the \Rclass{gAddChildMessage} class has two methods:
\Rfunction{initialize} and \Rfunction{handleMessage}.  The
\Rfunction{initialize} method properly fills the slots of the object when it
is created and the \Rfunction{handleMessage} method ensures that the new
model and \Robject{MVC} objects are created (which is what the
\Rfunction{handleMessage} method for the \Rclass{gAddDataMessage} class does)
and then it fills the \Rslot{linkData} slot of the new child model
so that the \Rfunction{toParent} and \Rfunction{fromParent} functions are
available for messaging between \Robject{MVC} objects and finally it stores
information in the child \Robject{MVC} object's controller that determines if
the child model was created from a subset of the parent
model.  Thus, the \Rfunction{handleMessage} method for the
\Rclass{gAddChildMessage} class both creates the child \Robject{MVC} object
and relates it to the parent \Robject{MVC} object.
